\subsection{Overview}
\begin{itemize}
    \item C is general purpose, you can use C to do anything you like, you can write OS programs.
    \item Uses statements that change a program's state.
    \item C is imperetive, it's going to focus how to get things done and using statements to do so.
    \item It's the most widely used language.
    \item C is a modern language:
        \begin{itemize}
            \item Basic control structures
            \item Top down planning (diferent steps can give you the same result, you don't have to follow the watefall aproach).
            \item Organized aroud the use of funtions
            \item Reliable, it won't crash for no reason, it's readable.
        \end{itemize}
    
    \item It's used in minicomputers, Unix/Linux for example.
    \item C is preferred language forproducing word processing programs, spreadsheets and compilers.
    \item C is popular for programming embedding systems, for example you see C being used in car components.
    \item C plays a strong role in the Linux operating system and the Unix base of MacOS.
    \item  C programs are easy to adapt to new models or languages.
    \item 1990, Software houses began turning to the C++ language for the large projects.
    \item C is a subset of C++ with object oriented programming tools added:
        \begin{itemize}
            \item Any C program is a valid C++ program, C is organized around functions and C++ is object oriented.
            \item By learning C you learn C++.
        \end{itemize}
    
    \item C remains a very demanding skill for professional applications.
    \item It's not an old language it's used, it's a valuable skill.
    \item C provides constructs that map efficiently:
        \begin{itemize}
            \item It provides low-level access to memory, low level capabilities.
            \item Requires minimal run-timesupport.
        \end{itemize}
    
    \item Invented by \textbf{Dennis Ritchie} of bell Laboratories, 1972; for the design of UNIX operating systems, C being old is an \textbf{advantage}.
    \item Was crated as a tool for working programmers, main goal is to be usefull language.
        \begin{itemize}
            \item No more will you need to get 20 lines of code to add two numbers.
            \item Realability and writability.
        \end{itemize}
    
    \item C initially became as the development language of the UNIX operating system, all mayor operating systems are written in C and/or C++.
    
    \item C evolved from a previous programming language named B:
        \begin{itemize}
            \item Uses many of the important concepts of B while adding data typing and other powerful features.
            \item B was a typeless language, every data item occupied one word in memory, and the burden of typing variables fell on the shoulders of the programmers.
            \item There is also a language called D.
            \item B didn't have data types.
        \end{itemize}
    
    \item C is available for most computers.
    \item C is hardware independent.
    \item 1970s C had evolved to traditional C.
    \item Variations of C, there are diferent variations of C:
        \begin{itemize}
            \item C89/C90: is supported by all current C compilers.
            \item C99: revises and and expends the capabilities of C, not all compilers suport it.
            \item C11: isn't suported by compilers and some features, focus on C99.
        \end{itemize}
    
    \item Standardizing C made the program portable and now any machine could run it.
    \item C is one of the most important languages.
    \item From C it's easy to migrate to any other programming language like Java, or objective-C.
\end{itemize}
