\subsection{Main features}
\begin{itemize}
    \item Efficiency and portability:
        \begin{itemize}
            \item Compact and fast, similar to assembly language.
            \item It is such a low level you can maximize efficiency.
            \item You can run it in several operating system.
            \item C is the leader in portability.
            \item Compilers are widely available.
            \item Linux and Unix come with a C compiler.
            \item C compilers exist in an array of forms, there are compilers for automobile C.
            
        \end{itemize}
    
    \item Powerful and flexible: 
        \begin{itemize}
            \item The unix/linux kernel is written in C, the kernel is the brain of the operating system.
            \item Many compilers and interpreters for other languages like FORTRA, Perl, Python, Pascal, LISP, Logo, Basic, etc.
            \item C programs are also used for solving physiscs and engeneering problems and for movie animations and special effects.
            \item C is flexible:
                \begin{itemize}
                    \item You can solve problems in diferent ways.
                    \item It's the basis for other programming languages.
                    \item It's a trade off. 
                \end{itemize}
            
            \item It is used for mobile phone apps, a form of objective C.
            \item Easy toi learn if it is your first language.
        \end{itemize}

    \item Programer oriented:
        \begin{itemize}
            \item Gives you access to memory and to hardware.
            \item You can manipulate individual bits in memory.
            \item Contains a large selection of operators.
            \item Less strict, this is an advantage and a disadvantage, you can make something much better but you need to be more responsable.
            \item Implementations have a large library of usefull C functions.
        \end{itemize}
    
    \item Other features:
        \begin{itemize}
            \item Provides bit manipulation.
            \item Provides features of lower level languages.
            \item You can access memroy by bits.
            \item You can have pointers, it's memory related.
        \end{itemize}
    
    \item Disadvantage:
        \begin{itemize}
            \item Flexibility and freedom also requires added flexibility, poiters especially, errors are dificult to trace.
            \item Sometimes because of it's wealth of operators it's dificult to read.
        \end{itemize}
    
    \item Advantages of C:
        \begin{enumerate}
            \item Fast and efficient
            \item Portable 
            \item Variety of data types and powerful operators
            \item Easy to extend
            \item Modularity 
            \item Function rich libraries
        \end{enumerate}
\end{itemize}
