\subsection{Four fundamental tasks in the creation of a C program}
\begin{enumerate}
    \item Editing: 
        \begin{itemize}
            \item Process of creating and modiffying your C source code.
            \item The source code is inside a file that contains the program.
            \item Choose a wise name for your base file name (.c).
            \item We'll use an IDE (code blocks) for the editing task. 
        \end{itemize}

    \item Compiling: 
        \begin{itemize}
            \item Converts the source code and turns it in to machine code and detects errors.
            \item The input to the compiler is going to be the source code.
            \item It's a two stage process
            \begin{enumerate}
                \item preprocessing fase (your code maybe modified or added to) 
                \item generation of the assembly language code.
            \end{enumerate} 
            \item The compiler will look at all the statements and shecks it to ensure that it confirms to the syntax and syntax.
            \item The compiler will detect dead code, typical errors such as syntax or semantic errors, but \textbf{not} the logic errors.
            \item The object code is a .o or .obj will be the assembly language statement into actual machine instructions.
            \item Run the command cc -c myprog.c of gcc -c myprog.c .
            \item If you omit this flag will automatically be linked as well.
        \end{itemize}
        
    \item Linking: 
        \begin{itemize}
            \item After the program has been compiled, linking is getting all the dependencies of the .o or .obj then it will take all of them and create a .exe or a.out file which is an executable.
            \item This fase will also detect errors.
            \item Non-existing library component for example will be an error.
            \item The program libraries suport and extend the C language by providing functionality already built and for me to use.
            \item If there is a linking failiure means you need to re-edit your source code.
            \item The programs that are big need to be modular.
        \end{itemize}

    \item Executing:
        \begin{itemize}
            \item Just run the program by clicking on it, or on the command line.
            \item This fase will be for executing your program.
            \item This fase also consists of debugging fase.
        \end{itemize}
\end{enumerate}

These processes are assentially the same to developing programs in any enviroment. 
