\begin{itemize}
    \item Notice the syntax: ``\#include <stdio.h>'', we know that this is a preprocesor directive.
    \item Directive = Statement
    \item Not strictly part of the program necesarily but the compiler will complain if you don't have it.
    \item These directives can be anywhere in the source code but it's convention to have them in the very first line.
    \item This statement says to the program to get the header files needed to compile correctly. The ``.h'' is a header files, they are called headers because they must be the first thing (by convention) in your source code.
    \item Header files:
        \begin{itemize}
            \item They define information about some of the funcitons used in the source code.
            \item The stdio.h is a C standard library file, for example ``printf()'', it comes with the C programming language.
            \item stdio.h is short for standard input/output, which means that we are going to use this library for input output operation.
            \item Header files specify infromation that the compiler uses to integrate any predefined functions within a program.
            \item We'll create our very own header files.
            \item It helps with the modularity feature of quality code.
        \end{itemize}
    
    \item Syntax of header files:
        \begin{itemize}
            \item Header files are case sensative, always use lower case.
            \item You can call header files in one of two ways:
                \begin{itemize}
                    \item With angle brackets: \#include <DavidsHeaderFile.h>; this syntax is used to call for the header specified in \textbf{one or more standard system directories}.
                    \item With double quotes: \#include "DavidsHeaderFile.h"; this double quote syntax tells C to look for the header file in the \textbf{current directory}.
                \end{itemize}
            
            \item The double quotes are usually for libraries made by developers.
            \item Every C compiler that conforms to the C11 standard will have a set of standard header files supplied with it.
            \item  
        \end{itemize}
    
    \item \#ifndef and \#define to protect against multiple inclusions of a header file; this is used to not import headers more than once.
\end{itemize}
