\section{Further topics to study}
\begin{itemize}
    \item More on data types:
        \begin{itemize}
            \item Defining your own types (typedef). Create your own data types.
        \end{itemize}
    
    \item More on the preprocessor:
        \begin{itemize}
            \item String concatenation. 
            \item You can optimize a lot of your code.
        \end{itemize}
    
    \item More on void*'s:
        \begin{itemize}
            \item A lot more.
        \end{itemize}
    
    \item Static libraries and shared objects.
        \begin{itemize}
            \item Reuse code.
        \end{itemize}

    \item macros
        \begin{itemize}
            \item Some of the functions we used are macros.
        \end{itemize}
    
    \item unions:
        \begin{itemize}
            \item Used in data structures.
        \end{itemize}
    
    \item Function pointers.
    \item Advanced pointers.
        \begin{itemize}
            \item Pointers to pointers.
        \end{itemize}
    
    \item Variable arguments to functions (variadic functions).
    \item Dynamic linking (dlm\_open).
        \begin{itemize}
            \item Open libraries on run-time.
        \end{itemize}
        
    \item Signals, forking and iner-process communications.
        \begin{itemize}
            \item Signals it's when two processes want to communicate, forking is creating another executing thread.
        \end{itemize}
    
    \item Threading and concurrency:
        \begin{itemize}
            \item Asynchronous execution.
        \end{itemize}
    
    \item Sockets:
        \begin{itemize}
            \item Data from computer to computer.
        \end{itemize}
    
    \item setjmp and longjmp for restoring state.
    \item More on memory management and fragmentation.
        \begin{itemize}
            \item Avoid fragmentation.
        \end{itemize}
    
    \item More and making your program portable. 
    \item Interfacing with kernel modules (drivers and ioctls).
        \begin{itemize}
            \item Create drivers for example.
        \end{itemize}

    \item More on compiler and linker flags.
    \item Advanced use of gdb:
        \begin{itemize}
            \item Debugging and more insight while using it.
            \item Understanding or interpreting the core files.
        \end{itemize}
    
    \item Profiling and tracing tools (gprof, dtrace, strace):
        \begin{itemize}
            \item Checking for dead code, tracing and other things.
        \end{itemize}
    
    \item Memory debugging tools such as valgrind.
        \begin{itemize}
            \item Find memory links and other memory related errors.
        \end{itemize}
\end{itemize}


%----------------------------------------------------------------------------------------
\section{Course summary}
\begin{itemize}
    \item Code Blocks: IDE.
    \item Basic operators: logical, arithmetic and assignment.
    \item Conditional statements: making decisions (if, switch).
    \item Repeating code: looping (for, while, do-while).
    \item Arrays: defining and initializing, multidimensional. 
    \item Functions: declaration and use, arguments and parameters, call by value vs. call by reference.
    \item Debugging: call stack, common mistakes, understanding compiler messages. 
    \item Structs: initializing, nested structures. 
    \item Character strings: basics, arrays of chars, character operations.
    \item Pointers: definition and use, using with functions and arrays, malloc, pointer arithmetic. 
    \item The preprocessor: \#define, \#include.
    \item Input and output: command line arguments scanf, printf. 
    \item File Input/Output: reading and writing to a file, fgetc, fgets, fputc, fgetc, fseek, etc. 
    \item Standard C library: string functions, math functions, utility functions, standard header files.
\end{itemize}

\subsection{Course outcomes}
\begin{itemize}
    \item You are now able to write beginner C programs.
    \item You are now able to write efficient C programs. 
        \begin{itemize}
            \item Modular.
            \item Low coupling.
            \item Naming conventions, indentation.
        \end{itemize}
    
    \item You are now able to find and fix errors:
        \begin{itemize}
            \item You now understand compiler messages. 
            \item You know how to use a debugger. 
        \end{itemize}
    
    \item You now understand fundamental aspects of the C programming language.
\end{itemize}
