\section{Polish \& Reverse polish notations}
\subsection{Polish notation}
\begin{itemize}
    \item In order to evaluate expressions. For example $(a+b)*c$. 
        \begin{itemize}
            \item We must first understand precedence, which means, the parenthesis are done first because it has a higher precedence, what ever is inside the parenthesis will be done first and then multiplied by $c$.
        \end{itemize}
    
    \item There are a handful of notations, the above example $(a+b)*c$ is called the infix notation. 
        \begin{itemize}
            \item There are other notations such as polish and reverse polish notation. 
        \end{itemize}
    \item In order to convert $a+b$  notation to polish notation:
        \begin{center}
            \begin{figure}[H]
                \begin{tikzpicture}[auto]
                    \node [] (1) at (0,0) {$a+b$}; 
                    \node [anchor=west] (2) at (2,0.75) {$+ab$ \quad prefix/polish}; 
                    \node [anchor=west] (3) at (2,-0.75) {$ab+$ \quad post-fix/reverse polish}; 

                    \path [line] (1) -- (2);
                    \path [line] (1) -- (3);
                \end{tikzpicture}
            \end{figure}
        \end{center}
    
    \item For evaluating infix expressions complexity tends to rise, since you need to keep details like what is being evaluated and so on, this complexity diminishes with the prefix and post fix notations. This is why we must aspire to make a converter of infix to prefix or post fix notation. 
        \begin{itemize}
            \item This is beneficial and reduces complexity. 
        \end{itemize}
    
    \item Example: given the infix expression $(a+b)*c$ find the equivalent in terms of a prefix expression.
        \[
          \overbrace{(a+b)}^{+ab}*c = \overbrace{+ab}^{r}*c = r * c = *rc = *+abc
        \]
        \begin{itemize}
            \item $*+abc$ this is going to be read from right to left, the first operator is going to be $+$ this allows precedence and allows for no use of braces. 
        \end{itemize}
\end{itemize}

\subsection{Reverse polish notation}
\begin{itemize}
    \item In order to convert from infix to polish notation using the example $(a+b)*c$: 
        \[
            \overbrace{(a+b)}^{ab+}*c = \overbrace{ab+}^{r}*c = r*c = rc* = ab+c*
        \]
    
    \item In the reverse polish notations reading it from left to right the first operation will be $+$. 
\end{itemize}


%----------------------------------------------------------------------------------------
\section{Understanding precedence of operators, conversion idea - infix to prefix/post-fix}
\begin{itemize}
    \item Let \$ be a substitute for the character \verb|^|, to represent exponential operations. 
    \item Remember the precedence of operators go as follows:
        \begin{center}
            \begin{tabular}{ |c|c| }
                \hline
                $\uparrow$  & () \\
                \hline
                $\uparrow$  & *, / \\
                \hline
                $\uparrow$  & +, - \\
                \hline
            \end{tabular}
        \end{center}

    
    \item Given the following expression: $a+b\$(c-d)k\$p$ convert to polish notation. 
        \begin{multline*}
            a+b\$(c-d)k\$p = a*b\$-cd/k\$p = a*\$-cd/k\$p = a*\overbrace{\$b-cd}^{r_1}/\overbrace{\$kp}^{r_2} \\ 
            = a*r_1/r_2 = *ar_1/r_2 = *ar_1/r_2 = /*ar_1r_2 = */a\$b-cd\$kp
        \end{multline*}
        \begin{itemize}
            \item The equivalent expression in prefix notation reading from right to left $*/a\$b-cd\$kp$
        \end{itemize}
    
    \item Given the same expression convert it to post-fix.
        \[
            a+b\$(c-d)k\$p = a*b\$cd-/k\$p = a*bcd-\$/kp\$ = abcd-\$*/kp\$ = abcd-\$*kp\$/
        \]
\end{itemize}


%----------------------------------------------------------------------------------------
\section{How to evaluate polish or reverse polish notation}
\begin{itemize}
    \item Iterating a string \verb|"24+32-*"| written in reverse polish notation. Program to implement has to iterate and will for this example only work for single digit numbers.
    \item Reverse polish notation, remember, is read from left to right.
        \begin{center}
            \begin{bytefield}{7}
                \bitheader{0-6} \\ 
                \bitbox{1}{2} & 
                \bitbox{1}{4} & 
                \bitbox{1}{+} & 
                \bitbox{1}{3} & 
                \bitbox{1}{2} & 
                \bitbox{1}{-} & 
                \bitbox{1}{*} 
            \end{bytefield}
        \end{center}
    
    \item Position 0 and 1 are operands and must be pushed to the stack. 
    \item Position 2 is, however, an operator, poping the stack twice will allow us to perform the operation of adding the operands, the result of both is pushed to the stack. 
    \item Position 3 and 4 are operands, the operator on position 5 pops the stack twice, subtracts the operands and pushes the result to the stack. 
    \item The position 6 operand pops the stack twice and performs the operation, pushing the value to the stack, poping the stack one time will reveal the final result. 
    \item For the polish notation start at the end and do the same thing. 
\end{itemize}


%----------------------------------------------------------------------------------------
\section{Algorithm for evaluating post-fix expression}
\begin{algorithm}[H]
    \SetAlgoLined
    \large
    initialize(stack)\;
    \While{not end of the post-fix string}{
        next\_current = gets the next token from the string\;
        \uIf{next\_token is an operand}{
            push(stack, next\_token)\;
        } \uElseIf{next\_token is an operator}{
            operand1 = pop(stack)\; 
            operand2 = pop(stack)\; 
            result = operate(operand2, operand1, next\_token)\; 
            push(stack, result)\; 
        }
    }
    print pop(stack)\; 
    end procedure interpret post-fix expressions\;
\caption{Interpret post-fix expressions pseudo-code}
\end{algorithm}

%----------------------------------------------------------------------------------------
\section{Implementing evaluation of post-fix expressions with the C programming language}
\inputcode{c}{\code/postfix_eval.c}



%----------------------------------------------------------------------------------------
\section{Understanding the precedence function}
\begin{itemize}
    \item The precedence function takes two arguments, the first argument is the stack top, the second is the token, the function returns true if the first argument (stack top) has equal or higher precedence than the token. 
    \item The function will return false if the token has higher precedence than the second.
    \item The following rules apply here: (op being any operator encountered)
        \begin{itemize}
            \item Any operator existing in the stack compared with the \verb|(| character will return false, meaning the \verb|(| has higher precedence than any other operator.
                \begin{minted}[autogobble]{c}
                    precedence(stack_top,'('); // false
                \end{minted}
            
            \item If the stack top is \verb|(| character, any operator stored in token has higher precedence, thus the function will return false.
                \begin{minted}[autogobble]{c}
                    precedence('(',token); // false
                \end{minted}
            
            \item If the precedence function has a token or second argument equal to the character \verb|)|, this means which ever character stored in the stack top has a higher precedence than closing parenthesis, except the case when the stack top is \verb|(| opening parenthesis.
                \begin{minted}[autogobble]{c}
                    precedence(stack_top,'('); // true
                    // except: stack_top = '('
                \end{minted}
            
            \item If the stack top is \verb|(| opening parenthesis, and the token is \verb|)| closing parenthesis, this is false.
                \begin{minted}[autogobble]{c}
                    precedence('(',')'); // false
                \end{minted}
        \end{itemize}
    
    \item Remember the order of operations or precedence is as follows in mathematics:
        \begin{enumerate}
            \item Anything within parenthesis.
            \item Exponentiation and root extraction. 
            \item Multiplication and division.
            \item Addition and subtraction.
        \end{enumerate}
    
    \item For all intents and purposes for this program if the precedence is the same the precedence function will return true as if one were greater than the other. The following examples apply:
        \input{figs/precedence_results.tex}
\end{itemize}
%----------------------------------------------------------------------------------------
\section{Explaining how it works}
\begin{itemize}
    \item Given the expression $m+(a+((b-c)*(d+k)))\$(x+y)*p$; its worth noting that this expression can be interpreted as follows:
        \begin{align*}
            =\;&m+\overbrace{(a+((b-c)*(d+k)))}^{r}\$\overbrace{(x+y)}^{w}*p \\    
            =\;&m+r^w*p \\ 
        \end{align*}
    
    \item Let's index the string:
        \begin{center}
            \begin{tabular}{ c }
                \mintinline{c}{infix_string =} \\ \\ \\ \\ 
            \end{tabular}
            \begin{bytefield}{27}
                \bitheader{0-26} \\ 
                \bitboxes{1}{{m} {+} {(} {a} {+} {(} {(} {b} {-} {c} {)} {*} {(} {d} {+} {k} {)} {)} {)} {\$} {(} {x} {+} {y} {)} {*} {p} } \\
            \end{bytefield}
        \end{center}
    
    \item We'll next iterate throughout the entire length of the string, using a loop and a counter variable.
\end{itemize}

\subsection{Each index in the example}
\input{figs/index_example.tex}


%----------------------------------------------------------------------------------------
\section{Writing the algorithm for converting infix expression to equivalent post-fix}
\begin{itemize}
    \item Arrays:
        \begin{itemize}
            \item infix: to hold the infix string. 
            \item Post-fix: to hold the post-fix string, initially empty.
        \end{itemize}
    
    \item Other data structures:
        \begin{itemize}
            \item OperatorStack: top is initialized and push and pop operation works on this. 
        \end{itemize}
\end{itemize}
Pseudo-code for the infix to post-fix: 
\begin{center}
    \begin{algorithm}[H]
        \SetAlgoLined
        % \DontPrintSemiColon
        % \Input{}
        % \Output{}
        % \BlankLine
        \large
        Initialize\;
        \While{not the end of the infix string}{
            token = get the next element from infix string\; 
            \uIf{token is an operand}{
                append the token with the postfix string\;
            }
            \uElseIf{token is an operator}{
                \While{not empty operator stack and prcd(stacktop,token)}{
                    top\_operator = pop(operatorstack)\; 
                    append top\_operator with the postfix string\;     
                }
                \eIf{token = ')'}{
                    pop(operatorstack)\; 

                }{
                    push(operatorstack,token)\;
                }
            }
        }
        \While{not empty operatorstack}{
            top\_operator = pop(operatorstack)\;
            append top\_operator with the postfix string\; 
        }
        print postfix\;
        \caption{Algorithm for converting infix to post-fix}
    \end{algorithm}
\end{center}


%----------------------------------------------------------------------------------------
\section{Combine the conversion and evaluation function in a single program}
\begin{itemize}
    \item Now we can combine the program that evaluates the post-fix notation and the infix to post-fix notation.
\end{itemize}
\inputcode{c}{\code/calculator.c}


